Binning Strategies and Related Loss for Large Data

### 2015-04-17 Paper.Rnw
clarification about "reduced binned data" to acknowlege that other bin summaries could be recorded, however in our paper we only consider frequency because it is the only thing required to construct a binned scatterplot

Data reduction comes at the expense of spatial information of any of the individual points. After aggregation the original spatial locations cannot be recovered. The loss of information incurred from binning will be explored in following sections.

@@

%-----------------------------------------------------------------------------
\subsection{Choice of bin width}

@@

The reduced binned data from spatial binning contains the center and count information for all bins. 
\hh{This sounds like it is set in stone - it's not. We can define which information is kept for each of the bins. Make this an assumption rather than a statement and define what kind of center you are talking about - visual or numeric. }

The bin frequencies may themselves be binned in order to discretize the color scale for the binned scatterplot.
The loss of frequency information in frequency binning is dependent on the selection of a discrete color mapping, where the binning algorithm and number of frequency bins must be specified. To explore frequency loss characteristics using the simulated bivariate data sets, each of the three fine resolution sets were first spatially binned with standard rectangular binning with 5 X 5 unit bins and origin at (0,0). Then standard and quantile frequency binning algorithms were applied to either the bin counts or log bin counts. As discussed in Section~\ref{FreqLoss}, the average person is able to distinguish about seven colors simultaneously, so we will limit the considered options to using between one and ten frequency bins, near the upper limit of human perceptual abilities \citep{colorperception}.

>>>>

Note that numerical attributes other than frequency of the binned data may also be recorded during binning, however only frequency is required to construct a binned scatterplot. Data reduction comes at the expense of spatial information of any of the individual points. After aggregation the original spatial locations cannot be recovered. The loss of information incurred from binning will be explored in following sections.

@@

%-----------------------------------------------------------------------------
%\subsection{Choice of bin width}

@@

The reduced binned data from spatial binning contains the center and count information for all bins. 
%\hh{This sounds like it is set in stone - it's not. We can define which information is kept for each of the bins. Make this an assumption rather than a statement and define what kind of center you are talking about - visual or numeric. }
The bin frequencies may themselves be binned in order to discretize the color scale for the binned scatterplot.
The loss of frequency information in frequency binning is dependent on the selection of a discrete color mapping, where the binning algorithm and number of frequency bins must be specified. To explore frequency loss characteristics using the simulated bivariate data sets, each of the three fine resolution sets were first spatially binned with standard rectangular binning with 5 X 5 unit bins and origin at (0,0). Then standard and quantile frequency binning algorithms were applied to either the bin counts or log bin counts. As discussed in Section~\ref{FreqLoss}, the average person is able to distinguish about seven colors simultaneously, so we will limit the considered options to using between one and ten frequency bins, near the upper limit of human perceptual abilities \citep{colorperception}.
